---
title: "extra"
author: "Rachel Holman"
date: "2023-07-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(GGally)
library(multcomp)
library(leaps)
library(car)
library(ROCR)
```



## extra model

```{r}
## Final model to really address question: do these variables impact odds of above average level of construction and design

final <- glm(highGrade~ price + bedrooms + bathrooms + sqft_living + 
                sqft_lot + floors + houseAge, 
              family="binomial", data=train)

summary(final)
```

Estimated logistic regression equation:
$$highGrade = -4.305 + 0.000006244(price) + -0.4280(bedrooms) + 0.2250(bathrooms) + \\ 0.001572(sqft\_living) + -0.000001888(sqft\_lot) + 0.5828(floors) + -0.03588(houseAge)$$


```{r}
##predicted probs for test data
preds<-predict(final,newdata=test, type="response")

##produce the numbers associated with classification table
rates<-ROCR::prediction(preds, test$highGrade)

##store the true positive and false positive rates
roc_result<-ROCR::performance(rates,measure="tpr", x.measure="fpr")

plot(roc_result, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red") 
```

```{r}
##confusion matrix with threshold of 0.5
table(test$highGrade, preds>0.5)

#0.5
fpr0.5=502/(2913+502)
tpr0.5=2545/(2545+524)
accuracy= (2913+2545)/(2913+502+2545+524)
accuracy

plot(roc_result, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red") 
points(x=fpr0.5, y=tpr0.5, col="blue", pch=16)
```

```{r}
auc<-performance(rates, measure = "auc")
auc@y.values
```

An AUC closer to 1 indicates the model does better than random guessing in classifying observations. Since it is almost equal to 1, it is classifying almost all observations correctly.


```{r}
table(test$highGrade, preds>0.5)
```

Error rate: $502+524 \over (2913+502+524+2545)$ =  $1026 \over 6484$ = 0.1582357  
False Positive Rate: $502 \over (2913+502)$ =  $502 \over 3415$ = 0.1469985    
False Negative Rate: $524 \over (524+2545)$ =  $524 \over 3069$ = 0.1707397    
True Positive Rate: $2913 \over (2913+502)$ =  $2913 \over 3415$ = 0.8530015    
True Negative Rate: $2545 \over (524+2545)$ =  $2545 \over 3069$ = 0.8292603  
Accuracy: $2913+2545 \over (2913+502+524+2545)$ =  $5458 \over 6484$ = 0.8417643  
Precision: $2545 \over (502+2545)$ = $2545 \over 3047$ = 0.8352478


```{r}
vif(final)
ggpairs(final, columns = c(1:8), progress=F)
```
Vif values all very small. Slightly high corrolation between sqft_living and price, and sqft_living and bathrooms. Nothing too concerning. 

## Hypothesis test: is this model useful?

$H_0 : \beta_{price} = \beta_{bedrooms} = \beta_{bathrooms} = \beta_{sqft\_living} = \beta_{sqft\_lot} = \beta_{floors} = \beta_{houseAge} = 0$ (The full model is not useful for estimating the grade status of a house)   
$H_A :$ At least one coefficient in $H_0$ is nonzero (The full model is useful for estimating the grade status of a heouse)   

```{r}
#likelihood ratio test
null <- glm(highGrade~1, family="binomial", data=train)

##test statistic
TS1<-null$deviance-final$deviance
TS1

##critical value
qchisq(1-0.05,7)

##pvalue
1-pchisq(TS1,7)
```

Test Statistic, $\triangle G^2$, = 10459.22  
Critical value = 14.06714  
P-value = 0

Because the test statistic is larger than the critical value, and the p-value is smaller than $\alpha=0.05$, we reject $H_0$ in favor of $H_A$. In other words, we have enough evidence to conclude that this logistic regression model with seven predictors is useful in estimating the grade status of a house compared to the intercept-only model.



```{r}
summary(final)
```

Estimated logistic regression equation:
$$highGrade = -4.305 + 0.000006244(price) + -0.4280(bedrooms) + 0.2250(bathrooms) + \\ 0.001572(sqft\_living) + -0.000001888(sqft\_lot) + 0.5828(floors) + -0.03588(houseAge)$$


$\beta_{price} = 0.000006244$. The estimated log odds of a house being ranked with above average condition and design increases by 0.000006244 for each additional dollar of a house is sold for, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(0.000006244) = 1.000006 for each additional dollar of a house is sold for, when controlling for the other predictors.    

$\beta_{bedrooms} = -0.4280$. The estimated log odds of a house being ranked with above average condition and design decreases by 0.4280 for each additional bedroom a house has, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(-0.4280) = 0.6518114 for each additional dollar of a house is sold for, when controlling for the other predictors. 

$\beta_{bathrooms} = 0.2250$. The estimated log odds of a house being ranked with above average condition and design increases by 0.2250 for each additional bathroom a house has, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(0.2250) = 1.252323 for each additional dollar of a house is sold for, when controlling for the other predictors.    

$\beta_{sqft\_living} = 0.001572$. The estimated log odds of a house being ranked with above average condition and design increases by 0.001572 for each additional square foot of interior living space a house has, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(0.001572) = 1.001573 for each additional dollar of a house is sold for, when controlling for the other predictors.   

$\beta_{sqft\_lot} = -0.000001888$. The estimated log odds of a house being ranked with above average condition and design decreases by 0.000001888 for each additional square foot of land space a house has, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(-0.000001888) = 0.9999981 for each additional dollar of a house is sold for, when controlling for the other predictors.   

$\beta_{floors} = 0.5828$. The estimated log odds of a house being ranked with above average condition and design increases by 0.5828 for each additional floor or level a house has, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(0.5828) = 1.791046 for each additional dollar of a house is sold for, when controlling for the other predictors.    

$\beta_{houseAge} = -0.03588$. The estimated log odds of a house being ranked with above average condition and design decreases by 0.03588 for each additional year old a house is, when controlling for the other predictors. In other words, the estimated odds of a house being ranked with above average condition and design is multiplied by exp(-0.03588) = 0.9647561 for each additional dollar of a house is sold for, when controlling for the other predictors.    


