---
editor_options:
  markdown:
    wrap: 72
output: pdf_document
---

# Metadata

    Course:   DS 5100
    Module:   11 R Programming 2
    Topic:    HW on Tidyverse
    Author:   R.C. Alvarado (adapted)
    Date:     7 July 2023 (revised)

# Student Info

    Name:  Rachel Holman
    Net ID: dnw9qk
    File GitHub URL: 
    https://github.com/rachel-holman/DS5100-dnw9qk/blob/main/lessons/M11/hw11.Rmd

# Instructions

In your **private course repo** use this notebook to write code that
performs the tasks below.

Save your notebook in the `M11` directory.

Remember to add and commit these files to your repo.

Then push your commits to your repo on GitHib.

Be sure to fill out the **Student Info** block above.

To submit your homework, save your results as a PDF and upload it to
GradeScope.

**TOTAL POINTS: 7**

# Overview

In this homework, you will work with [the Abalone
dataset](https://archive.ics.uci.edu/ml/datasets/Abalone) from the UCI
Machine Learning Repository.

To get started, download and import the `abalone.data` dataset from this
URL:

-   <https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data>

You can pass the URL directly to `read.csv()` and that there is no
header row.

Note: The instruction to print in the questions below can be
accomplished either through the `print()` function or by displaying a
value directly.

**TOTAL POINTS: 7**

# Tasks

## Task 0

(0 points)

Get the dataset.

```{r}
library(tidyverse)
abalone <- read.csv('https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data', 
                    header=F)
head(abalone)
```

## Task 1

(1 point)

Print the number of rows in the dataset.

```{r}
print(nrow(abalone))
```

## Task 2

(1 point)

The rightmost column is the number of rings. Print the maximum number of
rings

```{r}
print(max(abalone$V9))
```

## Task 3

(1 point)

The leftmost column is the gender with these values: `M`: male, `F`:
female, `I`: infant.

Apply the `filter()` function from tidyverse to select only rows where
gender is infant, and print the number of records.

```{r}
inf <- filter(abalone, abalone$V1 == "I")
print(nrow(inf))
```

## Task 4

(1 point)

Apply the `filter()` function from `tidyverse` to select only rows where
gender is infant or male, and print the number of records.

```{r}
inf_or_m <- filter(abalone, abalone$V1 == "I" | abalone$V1 == "M")
print(nrow(inf_or_m))
```

## Task 5

(1 point)

Call the `table()` function on the abalone genders to find out how many
of each gender are present.

Print the result.

```{r}
table(abalone$V1)
```

## Task 6

(1 point)

Compute the mean value of column 2 (V2) grouped by gender.

V2 is the longest shell measurement.

Requirements: use the `%>%` operator to chain commands, and the
`group_by()` and `summarize()` functions.

```{r}
meanvals <- abalone %>%
  group_by(V1) %>%
  summarize(mean_shell = mean(V2)) %>%
  rename("Gender" = "V1",
         "Mean Shell Measurement" = "mean_shell")
meanvals
```

# Task 7

(1 point)

Compute the MEDIAN value of longest shell measurement for only the
males.

Requirements: use the `%>%` operator to chain commands.

```{r}
meanvals_m <- abalone %>%
  filter(V1 == "M") %>%
  group_by(V1) %>%
  summarize(med_shell = median(V2)) %>%
  rename("Gender" = "V1",
         "Median Shell Measurement" = "med_shell")
meanvals_m
```
