---
title: "Practice Questions over R Basics"
author: "Rachel Holman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started with R

#### 1. The mtcars dataset comes built-in with R. The dataset was extracted from the 1974 Motor Trend US magazine and comprises several characteristics of automobile design and performance for 32 automobiles (1973 to 74 models).

**(a)** Read this data in and assign it to an object called cars.df
```{r}
cas.df <- mtcars
```


**(b)** Use the environment window and report how many observations and variables are in this dataset.

- *There are 32 observations and 11 variables in this data set.*



#### 2. Download the dataset “students.txt” from Collab. The dataset contains information on students taking an introductory statistics class at a large public university.

**(a)** Read this data in and assign it to an object called students.df
```{r}
students.df <- read.table("students.txt", header=TRUE)
```


**(b)** Use the environment window and report how many observations and variables are in this dataset.

- *There are 249 observations and 9 viariables in this data set.*



#### 3. Install the following packages:
(a) tidyverse
(b) faraway
(c) MASS
(d) leaps
(e) ROCR
(f) nycflights13
(g) gapminder
(h) palmerpenguins


```{r, eval=FALSE}
install.packages("tidyverse")
install.packages("faraway")
install.packages("MASS")
install.packages("leaps")
install.packages("ROCR")
install.packages("nycflights13")
install.packages("gapminder")
install.packages("palmerpenguins")
```



#### 4. Load the faraway package, and read in the dataset called cornnit and assign it to an object called corn.df

```{r}
library(faraway)
cord.df <- cornnit
```


---

## Data Types & Stuctures in R

#### 5. Are the following valid names for objects in R?

**(a)** 2020_Major

- *NO*, valid object names cannot begin with a number

**(b)** .2020.Age

- *NO*, if an object name begins with ".", it cannot be followed by a number

**(c)** #Courses.2020

- *NO*, valid object names cannot use "#"

**(d)** _courses_2020

- *NO*, valid object names cannot begin with "_"

**(e)** Fav_Sport20

- *YES*, this is a valid object name

**(f)** major 2020

- *NO*, valid object names cannot contain a space

**(g)** age(2020)

- *NO*, valid object names cannot use "(" or ")"

**(h)** FavSport_2020

- *YES*, this is a valid object name




#### 6. Create a numeric vector with the following 10 values: 13, 91, 36, 95, 9, 3, 61, 20, 22, 97

**Assign these 10 values to a vector called practice. Use R to find out if practice is a character, numeric, or logical type.***

```{r}
practice <- c(13, 91, 36, 95, 9, 3, 61, 20, 22, 97)
class(practice)
```




#### 7. For each of the following, determine if they are TRUE or FALSE. Try answering first without using R, then use R to confirm.

**(a)** practice[5]==5 *(FALSE)*
```{r}
practice[5]==5
```


**(b)** practice[10]!=97 *(FALSE)*
```{r}
practice[10]!=97
```


**(c)** (practice[1]+practice[2])<104 *(FALSE)*
```{r}
(practice[1]+practice[2])<104 
```


**(d)** (practice[1]+practice[2])<=104 *(TRUE)*
```{r}
(practice[1]+practice[2])<=104
```


**(e)** (practice[2]==91) & (practice[9]==22) *(TRUE)*
```{r}
(practice[2]==91) & (practice[9]==22)
```


**(f)** (practice[5]<9) | (practice[6]>=4) *(FALSE)*
```{r}
(practice[5]<9) | (practice[6]>=4)
```




#### 8. Create the following matrix in R and assign it to the object Mat.A.
$$\begin{bmatrix} 4 \  1 \ 3 \\ 6 \ 2 \ 1\end{bmatrix}$$

```{r}
Mat.A <- matrix(c(4,6,1,2,3,1),nrow=2,ncol=3)
Mat.A
```


**(a)** Give the following column names to Mat.A: “Huey”, “Dewey”, “Louie”.
```{r}
colnames(Mat.A) <- c("Huey", "Dewey", "Louie")
Mat.A
```


**(b)** Without using R, what is the output if we type Mat.A[2,1]?

- *6*

**(c)** Without using R, what is the output if we type dim(Mat.A)?

- *2 3*




#### 9. Convert the vector practice to a factor. What is the order of the levels in this factor?

```{r}
practice2<-factor(practice)
practice2
```

*Levels in order: 3, 9, 13, 20, 22, 36, 61, 91, 95, 97*

---

## R Markdown

#### 10. Type up your answers to these questions using R Markdown, and output an HTML file.



